### 2020-08-14 manuscript/manuscript.rmd
updated extended fig.3 based on most recent commit

```{r n_cat, fig.width=6, fig.height=12}
# Return the allocation midpoint, ie the percent of sites that are allocated to the
# 50% of the least frequent categories
allocation_midpoint = function ( pb_summary ) {
  x = pb_summary %>% pull( count ) %>% sort(decreasing = FALSE)
  x = x/ sum(x)
  y = cumsum( x )
  return ( round( median( y ) *100, 2 ) )
}
# Parse the last sample from chain 1 of each analysis
phil_cat_c1 =
  parse_phylobayes_last_sample("../trees_new/frequency/subsampled_Philippe2009_only_choanozoa.phy_Poisson_CAT_Chain_1.chain")
phil_cat60_c1 =
  parse_phylobayes_last_sample("../trees_new/frequency/subsampled_Philippe2009_only_choanozoa.phy_Poisson_nCAT60_Chain_1.chain")
whel_cat_c1 =
  parse_phylobayes_last_sample("../trees_new/frequency/subsampled_Whelan2017_strict.phy_Poisson_CAT_Chain1.chain")
whel_cat60_c1 =
  parse_phylobayes_last_sample("../trees_new/frequency/subsampled_Whelan2017_strict.phy_Poisson_CAT60_Chain_1.chain")
# Create a single tibble with summaries of all analyses
pb_summaries = list(
  summarise_sample( phil_cat_c1 ) %>%
    mutate ( chain=1, matrix="Philippe2009", model="Poisson+CAT"),
  summarise_sample( phil_cat60_c1 ) %>%
    mutate ( chain=1, matrix="Philippe2009", model="Poisson+nCAT60"),
  summarise_sample( whel_cat_c1 ) %>%
    mutate ( chain=1, matrix="Whelan2017_strict", model="Poisson+CAT"),
  summarise_sample( whel_cat60_c1 ) %>%
    mutate ( chain=1, matrix="Whelan2017_strict", model="Poisson+nCAT60")
) %>%
  bind_rows() %>%
  mutate( analysis = paste(matrix, model, sep=" ")) %>%
  mutate( model=factor(model, levels=c("Poisson+nCAT60", "Poisson+CAT")) )

@@

# Identify the midpoint of each set of allocations
mid_phil_cat =
  pb_summaries %>%
  filter( matrix=="Philippe2009" ) %>%
  filter( model=="Poisson+CAT" ) %>%
  allocation_midpoint()
mid_phil_ncat60 =
  pb_summaries %>%
  filter( matrix=="Philippe2009" ) %>%
  filter( model=="Poisson+nCAT60" ) %>%
  allocation_midpoint()
mid_whel_cat =
  pb_summaries %>%
  filter( matrix=="Whelan2017_strict" ) %>%
  filter( model=="Poisson+CAT" ) %>%
  allocation_midpoint()
mid_whel_ncat60 =
  pb_summaries %>%
  filter( matrix=="Whelan2017_strict" ) %>%
  filter( model=="Poisson+nCAT60" ) %>%
  allocation_midpoint()
n_categories_phil_cat =
  pb_summaries %>%
  filter( matrix=="Philippe2009" ) %>%
  filter( model=="Poisson+CAT" ) %>%
  nrow()
n_categories_whel_cat =
  pb_summaries %>%
  filter( matrix=="Whelan2017_strict" ) %>%
  filter( model=="Poisson+CAT" ) %>%
  nrow()
# Perform global MDS analysys
fit = cmdscale( dist( pb_frequencies ) ,eig=TRUE, k=2)
pb_summaries %<>% mutate( x_global = fit$points[,1], y_global = fit$points[,2] )
# Generate plots
gg_rank_phil =
  pb_summaries %>%
  filter( matrix=="Philippe2009" ) %>%
  ggplot() +
  geom_line(aes(x=rank, y=count, col=model)) +
  theme(legend.position="right") + ggtitle("Philippe2009_Choanimalia")
gg_rank_whel =
  pb_summaries %>%
  filter( matrix=="Whelan2017_strict" ) %>%
  ggplot() +
  geom_line(aes(x=rank, y=count, col=model)) +
  theme(legend.position="right") + ggtitle("Whelan2017_strict")
gg_mds_phil =
 pb_summaries %>%
  filter( matrix=="Philippe2009" ) %>%
  ggplot() +
  geom_point(aes(x=x_global, y=y_global, size=count, col=model), alpha=0.4 ) +
  theme(legend.position="right")
gg_mds_whel =
 pb_summaries %>%
  filter( matrix=="Whelan2017_strict" ) %>%
  ggplot() +
  geom_point(aes(x=x_global, y=y_global, size=count, col=model), alpha=0.4 ) +  
  theme(legend.position="right")
gg_chain =
ggarrange(
  gg_rank_phil,
  gg_rank_whel,
  gg_mds_phil,
  gg_mds_whel,
  ncol  = 2,
  nrow = 2,
  labels = c("A", "B", "C", "D"),
  common.legend = TRUE
)
These results further clarify when analyses support Porifera-sister (*e.g.* Philippe2009, Whelan2017). Outgroup sampling must be restricted (Choanimalia), a Poisson rather than GTR exchange matrix must be used, and a very large number of site categories must be used (unconstrained CAT, giving hundreds of equilibrium frequency categories). Analyses under other conditions either support Ctenophora-sister or are unresolved. This is very consistent across published analyses and our new panels of analyses. The question, then, isn't why similar analyses give different results, but how we should interpret variation in results when we run analyses that differ in these respects. If we fix the first two features to conditions that are necessary for Porifera-sister support (Choanimalia taxon sampling and a Poisson exchange matrix), there are several things we can learn from examining how the number of equilibrium frequency categories impacts results that sheds light on interpretation of those results.

>>>>

```{r n_cat}
# Return the allocation midpoint, ie the percent of sites that are allocated to the
# 50% of the least frequent categories
allocation_midpoint = function ( pb_summary ) {
  x = pb_summary %>% pull( count ) %>% sort(decreasing = FALSE)
  x = x/ sum(x)
  y = cumsum( x )
  return ( round( median( y ) *100, 2 ) )
}

# Parse the last sample from chain 1 of each analysis
phil_cat_c1 =
  parse_phylobayes_last_sample("../trees_new/frequency/subsampled_Philippe2009_only_choanozoa.phy_Poisson_CAT_Chain_1.chain")

phil_cat60_c1 =
  parse_phylobayes_last_sample("../trees_new/frequency/subsampled_Philippe2009_only_choanozoa.phy_Poisson_nCAT60_Chain_1.chain")

whel_cat_c1 =
  parse_phylobayes_last_sample("../trees_new/frequency/subsampled_Whelan2017_strict.phy_Poisson_CAT_Chain1.chain")

whel_cat60_c1 =
  parse_phylobayes_last_sample("../trees_new/frequency/subsampled_Whelan2017_strict.phy_Poisson_CAT60_Chain_1.chain")

# Create a single tibble with summaries of all analyses
pb_summaries = list(
  summarise_sample( phil_cat_c1 ) %>%
    mutate ( chain=1, matrix="Philippe2009", model="Poisson+CAT"),
  summarise_sample( phil_cat60_c1 ) %>%
    mutate ( chain=1, matrix="Philippe2009", model="Poisson+nCAT60"),
  summarise_sample( whel_cat_c1 ) %>%
    mutate ( chain=1, matrix="Whelan2017_strict", model="Poisson+CAT"),
  summarise_sample( whel_cat60_c1 ) %>%
    mutate ( chain=1, matrix="Whelan2017_strict", model="Poisson+nCAT60")
) %>%
  bind_rows() %>%
  mutate( analysis = paste(matrix, model, sep=" ")) %>%
  mutate( model=factor(model, levels=c("Poisson+nCAT60", "Poisson+CAT")) )

@@

# Identify the midpoint of each set of allocations
mid_phil_cat =
  pb_summaries %>%
  filter( matrix=="Philippe2009" ) %>%
  filter( model=="Poisson+CAT" ) %>%
  allocation_midpoint()

mid_phil_ncat60 =
  pb_summaries %>%
  filter( matrix=="Philippe2009" ) %>%
  filter( model=="Poisson+nCAT60" ) %>%
  allocation_midpoint()

mid_whel_cat =
  pb_summaries %>%
  filter( matrix=="Whelan2017_strict" ) %>%
  filter( model=="Poisson+CAT" ) %>%
  allocation_midpoint()

mid_whel_ncat60 =
  pb_summaries %>%
  filter( matrix=="Whelan2017_strict" ) %>%
  filter( model=="Poisson+nCAT60" ) %>%
  allocation_midpoint()

n_categories_phil_cat =
  pb_summaries %>%
  filter( matrix=="Philippe2009" ) %>%
  filter( model=="Poisson+CAT" ) %>%
  nrow()

n_categories_whel_cat =
  pb_summaries %>%
  filter( matrix=="Whelan2017_strict" ) %>%
  filter( model=="Poisson+CAT" ) %>%
  nrow()

# Perform global MDS analysys
fit = cmdscale( dist( pb_frequencies ) ,eig=TRUE, k=2)
pb_summaries %<>% mutate( x_global = fit$points[,1], y_global = fit$points[,2] )

# Generate plots
gg_rank_phil =
  pb_summaries %>%
  filter( matrix=="Philippe2009" ) %>%
  ggplot() +
  geom_line(aes(x=rank, y=count, col=model)) +
  xlab( "category rank" ) +
  ylab( "site count" ) +
  theme(legend.position="right") + ggtitle("Philippe2009_Choanimalia")

gg_rank_whel =
  pb_summaries %>%
  filter( matrix=="Whelan2017_strict" ) %>%
  ggplot() +
  geom_line(aes(x=rank, y=count, col=model)) +
  xlab( "category rank" ) +
  ylab( "site count" ) +
  theme(legend.position="right")+ ggtitle("Whelan2017_strict")

gg_mds_phil =
  pb_summaries %>%
  filter( matrix=="Philippe2009" ) %>%
  ggplot() +
  geom_point(aes(x=x_global, y=y_global, size=count, col=model), alpha=0.4 ) +
  xlim( -0.5, 0.5) +
  ylim( -0.6, 0.6) +
  xlab( "MDS1" ) +
  ylab( "MDS2" ) +
  theme(legend.position="right")

gg_mds_whel =
  pb_summaries %>%
  filter( matrix=="Whelan2017_strict" ) %>%
  ggplot() +
  geom_point(aes(x=x_global, y=y_global, size=count, col=model), alpha=0.4 ) +  
  xlim( -0.5, 0.5) +
  ylim( -0.6, 0.6) +
  xlab( "MDS1" ) +
  ylab( "MDS2" ) +
  theme(legend.position="right")

gg_chain =
ggarrange(
  gg_rank_phil,
  gg_rank_whel,
  gg_mds_phil,
  gg_mds_whel,
  ncol  = 2,
  nrow = 2,
  labels = c("A", "B", "C", "D"),
  common.legend = TRUE
)
```
These results further clarify when analyses support Porifera-sister (*e.g.* Philippe2009, Whelan2017). Outgroup sampling must be restricted (Choanimalia), a Poisson rather than GTR exchange matrix must be used, and a very large number of site categories must be used (unconstrained CAT, giving hundreds of equilibrium frequency categories). Analyses under other conditions either support Ctenophora-sister or are unresolved. This is very consistent across published analyses and our new panels of analyses. The question, then, isn't why similar analyses give different results, but how we should interpret variation in results when we run analyses that differ in these respects. If we fix the first two features to conditions that are necessary for Porifera-sister support (Choanimalia taxon sampling and a Poisson exchange matrix), there are several things we can learn from examining how the number of equilibrium frequency categories impacts results that sheds light on interpretation of those results.
